---
title: "Brazilian Vaccine Data"
author: "Paul Truche"
date: "11/16/2020"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#tinytex::install_tinytex()
options(scipen = 999)
```

```{r, include=FALSE, warning=FALSE}
library(readxl)
imunizacoes_2019 <- read_excel("imunizacoes 2019.xlsx")

imunizacoes_2019$Date<-as.Date(imunizacoes_2019$Date,
                        format = "%y/%m/%d")

All_data <- read_excel("All data.xlsx", sheet=1)
All_data$Date<-as.Date(All_data$Date,
                        format = "%y/d%/%m")

lessthanoneyear <- read_excel("All data.xlsx", sheet=2)
lessthanoneyear$Date<-as.Date(lessthanoneyear$Date,
                        format = "%y/d%/%m")

df1to4 <- read_excel("All data.xlsx", sheet=3)
df1to4$Date<-as.Date(df1to4$Date,
                        format = "%y/d%/%m")

df5to10 <- read_excel("All data.xlsx", sheet=4)
df5to10$Date<-as.Date(df5to10$Date,
                        format = "%y/d%/%m")

df10to14 <- read_excel("All data.xlsx", sheet=5)
df10to14$Date<-as.Date(df10to14$Date,
                        format = "%y/d%/%m")

df15to29 <- read_excel("All data.xlsx", sheet=6)
df15to29$Date<-as.Date(df15to29$Date,
                        format = "%y/d%/%m")

df30to49 <- read_excel("All data.xlsx", sheet=7)
df30to49$Date<-as.Date(df30to49$Date,
                        format = "%y/d%/%m")

df50to64 <- read_excel("All data.xlsx", sheet=8)
df50to64$Date<-as.Date(df50to64$Date,
                        format = "%y/d%/%m")

df65plus <- read_excel("All data.xlsx", sheet=9)
df65plus$Date<-as.Date(df65plus$Date,
                        format = "%y/d%/%m")
```

```{r, include=FALSE, warning=FALSE}
library(qicharts)
library(wpp2019)
library(hrbrthemes)
library(ggplot2)
library(ggthemes)
library(data.table)
library(tidyverse)
library(scales)
library(plyr)
library(reshape)
```

In this study, we examine how the number of vaccines adminstered in Brazil has changed as a result of the COVID19 pandemic.
<br><br><br><br>

Our data comes from DATASUS which tabulates the number of vaccines distributed by state for each month.  Here we utilize data from 2015 thorugh 2020.
<br><br><br><br>

```{r, echo=FALSE, warning=FALSE,fig.cap="Total vacccines adminstered across Brazil"}
df<-data.frame(imunizacoes_2019)

# Most basic bubble plot
ggplot(All_data, aes(x=Date, y=Total)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month") +
  ggtitle("Total Vaccines 2015:2020") +
  theme_fivethirtyeight()+
   scale_y_continuous(limits = c(3000000, 18000000)) +
  scale_color_fivethirtyeight()+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))
```
<br><br><br><br>

```{r, echo=FALSE, warning=FALSE}
dataforlong<-df[,c(1,3:7)]

dataforlong<-dataforlong %>% 
  dplyr::rename( North = North.vaccines, Northeast = Northeast.vaccines, Southeast =Southeast.vaccines , South =South.vaccines, Midwest =Midwest.vaccines)

dflong <- reshape::melt(dataforlong, id.vars="Date")

dflong<-dflong %>% 
  dplyr::rename(
    State = variable,
     Month = Date)

ggplot(dflong, aes(x=Month, y=value, group=State, color=State)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month") +
  ggtitle("Total By State 2015:2020") +
  theme_fivethirtyeight()+
  # scale_y_continuous(limits = c(0, 3000000)) +
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-9-1")))
```
<br><br><br><br>

Here we use an I chart followed by a moving range (MR) chart

```{r, echo=FALSE, warning=FALSE}
df$Total2<-df$Total.vaccines/(df$Total.pop/1000)
# Create vector of random values to plot

nationaltotal<-qic(Total2,
    x     = Date,
    data  = df,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 population)',
    xlab  = '',
    nint=30,
     x.format = "%b-%Y", main  = "Total Vaccinations Brazil: 2015-2020",sub ="")

nationaltotal<-qic(Total2,
    x     = Date,
    data  = df,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 population)',
    xlab  = '',
    nint=20,
     x.format = "%b-%Y", main  = "Total Vaccinations Brazil: 2015-2020",sub ="Break at March 2020", breaks=63)

qic(df$Total2,
    chart = 'mr',
    main  = 'Pairwise differences in Vaccine Rates (MR chart)',
    ylab  = 'Vaccines (Per 1000 pop)',
    xlab  = 'Month')
```
<br><br><br><br>
Below we use an xbar chart in order to incorperate vaccine rates per 1000 people by state.  
<br><br><br><br>

```{r, echo=FALSE, warning=FALSE}

df$Northadj<-df$North.vaccines/(df$North.pop/1000)
df$Northeastadj<-df$Northeast.vaccines/(df$Northeast.pop/1000)
df$Southeastadj<-df$Southeast.vaccines/(df$Southeast.pop/1000)
df$Southadj<-df$South.vaccines/(df$South.pop/1000)
df$Midwestadj<-df$Midwest.vaccines/(df$Midwest.pop/1000)

dataforlong2<-df[,c(1,15:19)]

dataforlong2<-dataforlong2 %>% 
  dplyr::rename( North = Northadj, Northeast = Northeastadj, Southeast =Southeastadj , South =Southadj, Midwest = Midwestadj)

dataforlong2 <- reshape::melt(dataforlong2, id.vars="Date")

dataforlong2<-dataforlong2 %>% 
  dplyr::rename(
    State = variable,
     Month = Date)

qic(value, 
    x     = Month, 
    data  = dataforlong2,
    chart = "xbar",
    main  = 'Average vaccination rate (Xbar chart; across states)',
    ylab  = 'Vaccines rate (per 1000 people)',
    xlab  = 'Month',x.format = "%b-%Y")

qic(value, 
    x     = Month, 
    data  = dataforlong2,
    chart = 'xbar',
    main  = 'Average vaccination rate (Xbar chart; across states)',
    ylab  = 'Vaccines rate (per 1000 people)',
     sub = "break at March 2020",
    xlab  = 'Month',breaks = 63,x.format = "%b-%Y")

```

```{r, echo=FALSE, warning=FALSE}
columns<-rownames(cbind(sort(colSums(All_data[,-c(1,2,3)]), T)[1:5]))

lessthanoneyearsubselected<-as.data.frame(All_data[,c("Date", "pop" ,columns)])

lessthanoneyearsubselected2 <- reshape::melt(lessthanoneyearsubselected, id.vars=c("Date", "pop"))

lessthanoneyearsubselected2$value<-lessthanoneyearsubselected2$value/(lessthanoneyearsubselected2$pop/1000)

ggplot(lessthanoneyearsubselected2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle("Vaccines by type (per 1000 persons") +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))
```

We can examine totals and individuals by age group
```{r, echo=FALSE, warning=FALSE}
#lessthanoneyear 
columns<-rownames(cbind(sort(colSums(lessthanoneyear[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(lessthanoneyear[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 births)")+
  ggtitle("Top 5 Vaccines age 0-1") +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-lessthanoneyear
qicchart$Total2<-qicchart$Total/(qicchart$pop)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "total vaccines less than one year of age",
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```

```{r, echo=FALSE, warning=FALSE}
#df1to4 
title<-"Top 5 Vaccines age 1-4"
columns<-rownames(cbind(sort(colSums(df1to4[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df1to4[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df1to4
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```


```{r, echo=FALSE, warning=FALSE}
#df5to10 
title<-"Top 5 Vaccines age 5-10"
columns<-rownames(cbind(sort(colSums(df5to10[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df5to10[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df5to10
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```


```{r, echo=FALSE, warning=FALSE}
#df10to14 
title<-"Top 5 Vaccines age 10-14"
columns<-rownames(cbind(sort(colSums(df10to14[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df10to14[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df10to14
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```


```{r, echo=FALSE, warning=FALSE}
#df15to29 
title<-"Top 5 Vaccines age 15-29"
columns<-rownames(cbind(sort(colSums(df15to29[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df15to29[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df15to29
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```


```{r, echo=FALSE, warning=FALSE}
#df30to49 
title<-"Top 5 Vaccines age 39-49"
columns<-rownames(cbind(sort(colSums(df30to49[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df30to49[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df30to49
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```


```{r, echo=FALSE, warning=FALSE}
#df50to64 
title<-"Top 5 Vaccines age 50-64"
columns<-rownames(cbind(sort(colSums(df50to64[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df50to64[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df50to64
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```


```{r, echo=FALSE, warning=FALSE}
#df65plus 
title<-"Top 5 Vaccines age 65+"
columns<-rownames(cbind(sort(colSums(df65plus[,-c(1,2,3)]), T)[1:5]))

subset<-as.data.frame(df65plus[,c("Date","pop", columns)])

subset2 <- reshape::melt(subset, id.vars=c("Date","pop"))

subset2$value<-subset2$value/(subset2$pop/1000)


ggplot(subset2, aes(x=Date, y=value, group=variable,color=variable)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month")  + ylab("number of vaccinations (per 1000 people)")+
  ggtitle(title) +
  theme_clean()+ theme(legend.title = element_blank(),legend.position="bottom",legend.background = element_rect(fill="white", 
                                  size=0.5, linetype="blank"))+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-10-1")))


qicchart<-df65plus
qicchart$Total2<-qicchart$Total/(qicchart$pop/1000)

qic(Total2,
    x     = Date,
    data  = qicchart,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 people)',
    xlab  = '',
    main = title,
    nint=30,
     x.format = "%b-%Y",sub ="Break at March 2020", breaks=63)
```

Here we examine the number of vaccines needed before age 1
```{r, echo=FALSE, warning=FALSE}
infantdf<-lessthanoneyear %>% select("Date","pop",
"BCG (BCG)",
"Hepatite B (HB)",
"Pentavalente (DTP+HB+Hib) (PENTA)",
"Poliomielite inativada (VIP)",
"Pneumocócica 10valente",
"Rotavírus pentavalente",
"Meningocócica Conjugada - C (MncC)",
"Influenza (INF)",
"Febre Amarela (FA)",
"Tríplice Viral (SCR)")

infantdf$infanttotal<-rowSums(infantdf[,2:12])

infantdf<-infantdf[!(infantdf$Date=="2020-11-01"),]

infantdf<-infantdf[-c(1:12),]
```


```{r, echo=FALSE, warning=FALSE}
#qic chart for each
infantdf$adj<-infantdf$`BCG (BCG)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "BCG (BCG)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Hepatite B (HB)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Hepatite B (HB)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Pentavalente (DTP+HB+Hib) (PENTA)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Pentavalente (DTP+HB+Hib) (PENTA)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Poliomielite inativada (VIP)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Poliomielite inativada (VIP)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Pneumocócica 10valente`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Pneumocócica 10valente",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Rotavírus pentavalente`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Rotavírus pentavalente",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Meningocócica Conjugada - C (MncC)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Meningocócica Conjugada - C (MncC)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Influenza (INF)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Influenza (INF)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Febre Amarela (FA)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Febre Amarela (FA)",
    nint=52,
     x.format = "%b-%Y",sub ="")

infantdf$adj<-infantdf$`Tríplice Viral (SCR)`/(infantdf$pop)

qic(adj,
    x     = Date,
    data  = infantdf,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 births)',
    xlab  = '',
    main = "Tríplice Viral (SCR)",
    nint=52,
     x.format = "%b-%Y",sub ="")
```


